Symmetric Circuits for Rank Gates

### 2017-12-16 sections/translating-formulas-to-cicuits.tex
editing process

  We note that in every iteration of the loop we have that every non-symmetric
  gate in $G^{\leq p}$ has unique labels. Thus, from Proposition
  \ref{prop:unique-labels-syntactic-equiv}, we may compute the syntactic
  equvilence relation on this set of gates efficiently. Thus it is easy to see
  that the above algorithm can be implemented so as to run in polynomial time.
\end{proof}

@@

\begin{lem}
  Let $C_n$ be a $(\mathbb{B}, \tau)$-circuit with unique labels of size $n$ and
  $\sigma \in \sym_n$. There is an algorithm takes as input a circuit $C_n$ and
  $\sigma \in \sym_n$ and outputs for each gate $g$ its image under the unique
  automorphism extending $\sigma$ (if it exists). This algorithm runs in time
  polynomial in the combined size of the input circuit and the encoding of the
  permutation.
  \label{lem:compute-automorphisms}
\end{lem}
\begin{proof}
  Suppose $C_n = \langle G, \Omega, \Sigma, \Lambda, L \rangle$. We build up the
  mapping $\pi$ extending $\sigma$ recursively. If at some point in the
  recursive construction we arrive at a point where no mapping for $g$ can be
  found we halt at that point and return that no automorphism exists.

@@

This result uses a proof very similar to the one for a similar result proved by
Anderson and Dawar~\cite{Anderson17D}.

\begin{lem}
  There is an algorithm that takes as input a circuit $C_n$ with unique labels
  and decides if $C_n$ is symmetric. If $C_n$ is symmetric outputs the orbits
  and coarsest supporting partitions of every gate. Moreover, this algorithm
  runs in time polynomial in the size of the input circuit.
  \label{lem:computing support}
\end{lem}

\begin{proof}
  Suppose $C_n = \langle G, \Omega, \Sigma, \Lambda, L \rangle$. Let $(u, v) \in
  \sym_n$ be a transposition. We may use Lemma \ref{lem:compute-automorphisms}
  to determine the unique automorphism $\pi_{(u,v)}$ extending $(u, v)$. It is
  easy to see that, since the set of transpositions generate $\sym_n$, for all
  $(u,v) \in \sym_n$ $\pi_{(u,v)}$ the automorphism $\pi_{(u,v)}$ exists if, and
  only if, $C_n$ is symmetric. We note that there are ${n}\choose{2}$
  transpositions. It follows that we may check in polynomial time if $C_n$ is
  symmetric. If $C_n$ is not symmetric we output that it is not symmetric and
  halt.

  Suppose then that $C_n$ is symmetric. For transposition $(u, v) \in \sym_n$,
  let $\mathcal{P}_{(u,v)} := \{\{u,v\}\}\bigcup_{w \in [n] \setminus
    \{u,v\}}\{\{w\}\}$ be a partition of $[n]$. We have that $\pi_{(u,v)}$ fixes
  $g$ if, and only if, $\mathcal{P}_{(u,v)}$ supports $g$. Since we can
  efficiently check if $\pi_{(u,v)}$ fixes $g$ it follows that we can
  effeciently check if $\mathcal{P}_{(u,v)}$ supports $g$. Let $\mathcal{P}$ be
  the result of combining the set consisting of those partitions
  $\mathcal{P}_{(u,v)}$ that support $g$ using $\mathcal{E}$. From Proposition
  \ref{prop:combining-supporting-patitions} we have that $\mathcal{P}$ supports
  $g$. Suppose that $\mathcal{P}$ is not the coarsest supporting partition of
  $g$. Then there exists a partition $\mathcal{P}'$ supporting $g$ such
  $\mathcal{P}' \preceq P$ and $\mathcal{P}' \neq \mathcal{P}$. But then there
  exists $P \in \mathcal{P}$ and $P' \in \mathcal{P}'$ such that $P \subsetneq
  P'$. But then there exists $a , b \in P'$ such that $a \not\in P$. It follows
  that the transposition $(a,b)$ fixes $\mathcal{P}'$ and so $\pi_{(a,b)}$ fixes
  $g$, but then, $\mathcal{P}_{(a,b)}$ supports $g$ and so, from the
  construction of $\mathcal{P}$, $\mathcal{P}$ is fixed by $(a,b)$. But
  $\mathcal{P}$ is not fixed by $(a,b)$, so we have a contradiction. We thus
  have that $\mathcal{P}$ is the coarsest supporting partition of $g$.

  It remains to compute the orbit of a gate $g$. Let $S_0 := \{g\}$ and for each
  $i \geq 0$ let $S_{i+1} := S_i \bigcup_{(u,v) \in \sym_n} \pi_{(u,v)}S_i$. Let
  $S$ be the least fixed point. It is easy to see that $S \subseteq \orb(g)$ as
  every gate in $S$ is equal to the action of some sequence of automorphisms
  (extending transpositions) acting on $g$. Moreover, since the set of all
  transpositions generates $\sym_n$, it follows that $\orb(g) \subseteq S$ and
  so $\orb(g) = S$.

@@

>>>>

  We note that in every iteration of the loop we have that every non-symmetric
  gate in $G^{\leq p}$ has unique labels. Thus, from Proposition
  \ref{prop:unique-labels-syntactic-equiv}, we may compute the syntactic
  equivalence relation on this set of gates efficiently. Thus it is easy to see
  that the above algorithm can be implemented so as to run in polynomial time.
\end{proof}

@@

\begin{lem}
  If $G \subseteq \sym_n$
  Let $C_n$ be a $(\mathbb{B}, \tau)$-circuit with unique labels of size $n$ and
  $\sigma \in \sym_n$. There is an algorithm takes as input a circuit $C_n$ and
  $\sigma \in \sym_n$ and outputs for each gate $g$ its image under the unique
  automorphism extending $\sigma$ (if it exists). This algorithm runs in time
  polynomial in the combined size of the input circuit and the encoding of the
  permutation.
  \label{lem:compute-automorphisms}
\end{lem}
\begin{proof}
  Suppose $C_n = \langle G, \Omega, \Sigma, \Lambda, L \rangle$. We build up the
  mapping $\pi$ extending $\sigma$ recursively. If at some point in the
  recursive construction we arrive at a point where no mapping for $g$ can be
  found we halt at that point and return that no automorphism exists.

@@

% \begin{lem}
%   Let $C_n$ be a $(\mathbb{B}, \tau)$-circuit with unique labels of size $n$ and
%   $\sigma \in \sym_n$. There is an algorithm takes as input a circuit $C_n$ and
%   $\sigma \in \sym_n$ and outputs for each gate $g$ its image under the unique
%   automorphism extending $\sigma$ (if it exists). This algorithm runs in time
%   polynomial in the combined size of the input circuit and the encoding of the
%   permutation.
%   \label{lem:compute-automorphisms}
% \end{lem}
% \begin{proof}
%   Suppose $C_n = \langle G, \Omega, \Sigma, \Lambda, L \rangle$. We build up the
%   mapping $\pi$ extending $\sigma$ recursively. If at some point in the
%   recursive construction we arrive at a point where no mapping for $g$ can be
%   found we halt at that point and return that no automorphism exists.

%   Let $h$ be any gate in the circuit. Suppose $h$ is an input gate. If $h$ is a
%   constant gate then let $\pi (h) = h'$. If $h$ is a relational gate such that
%   $R := \Sigma(h)$ and check if there exists $h'$ such that $\Sigma (h') = R$
%   and $\sigma \Lambda_R(h) = \Lambda_R(h')$ and, if $h$ is an output gate then
%   $h'$ is an output gate such that $\sigma Omega^{-1}(h) = \Omega^{-1}(h')$. If
%   such a $h'$ does not exist halt and output that there is no automorphism
%   extending $\sigma$. We have from Lemma \ref{} that if such an $h'$ exists it
%   must be unique, and we set $\pi(h) = h'$.

%   Let $h$ be an internal gate in the circuit and assume we have defined $\pi
%   (g)$ for every gate $g$ of height less than $h$. Let $h'$ be a gate in the
%   circuit such that $\Sigma(h) = \Sigma (h')$, $\pi L(h)$ is isomorphism
%   equivalent to $L(h')$ and, if $h$ is an output gate then $h'$ is an output
%   gate such that $\sigma \Omega^{-1}(h) = \Omega^{-1}(h')$. Since $C_n$ has
%   unique labels we have that $\pi L(h)$ is isomorphism equivalent to $L(h')$ if,
%   and only if, $L(h')^{-1}\pi L(h)$ acts like an isomorphism. This can be
%   determined easily. If no such $h'$ exists halt and output that there is no
%   automorphism extending $\sigma$. We have from Lemma \ref{} that if $C_n$ has
%   unique labels then there is at most one such $h'$.

%   This recursive approach can be implemented as an algorithm that runs in time
%   polynomial in the combined size of the inputs and outputs the required
%   automorphism if it exists.
% \end{proof}

This result uses a proof very similar to the one for a similar result proved by
Anderson and Dawar~\cite{AndersonD17}.

\begin{lem}
\label{lem:computing-support}
  Let $G \leq \sym_n$ be a subgroup and be $X$ a finite set. Suppose there there is an algorithm that $(g, x) \mapsto g \cdot x$ is computable in time polynomial in $\vert X \vert + n$. Suppose  $s \subset [n]$ be a support of $g$ There is an algorithm that takes in $x \in X$ and outputs $\orb_G (x)$ and $\SP_G(x)$. This algorithm runs in time polynomial in $\vert X \vert + n$.
\end{lem}
\begin{proof}
  Let $(u, v) \in \sym([n] \setminus \consp(G))$ be a transposition. We note that $(u,v) \in \stab(\consp(G))$ and there are ${n - \vert \consp(G) \vert}\choose{2}$ (and so polynomially) many such transpositions. For a transposition $(u, v) \in \sym ([n] \setminus \consp(G))$, let
  \begin{align*}
  \mathcal{P}_{(u,v)} := \{ \{u,v\}\} \bigcup_{w \in [n] \setminus \{ u,v \}} \{ \{ w \} \}
  \end{align*}
   be a partition of $[n]$. Then $\mathcal{P}_{(u,v)}$ supports $\stab_{\consp(G)}(x)$ if, and only if, $(u,v) \cdot x = x$. Then let $\mathcal{P}$ be the partition formed by applying $\mathcal{E}$ to all $\mathcal{P}_{(u,v)}$ such that $(u,v) \cdot x = x$ (which can be checked in polynomial time). From Proposition~\ref{prop:combining-supporting-patitions} we have that $\mathcal{P}$ supports $\stab_{\consp(G)}(x)$. Suppose that $\mathcal{P}$ is not the coarsest supporting partition of $\stab_{\consp(G)}(x)$. Then there exists a partition $\mathcal{P}'$ supporting $\stab_{\consp(G)}(x)$ such that $\mathcal{P}' \preceq P$ and $\mathcal{P}' \neq \mathcal{P}$. And so there exists $P \in \mathcal{P}$ and $P' \in \mathcal{P}'$ such that $P \subsetneq P'$. But then there exists $a , b \in P'$ such that $a \not\in P$. Note that It follows that the transposition $(a,b)$ fixes $\mathcal{P}'$ and so $(a,b) \in \stab_{\consp(G)}(x)$ (and so $a, b \not\in [n] \setminus \consp(G)$). But then $\mathcal{P}_{(a,b)}$ supports $\stab_{\consp(G)}(x)$ and so, from the construction of $\mathcal{P}$, $\mathcal{P}$ is fixed by $(a,b)$. But we selected $a$ and $b$ such that $\mathcal{P}$ is not fixed by $(a,b)$, and so we have a contradiction. We thus have that $\mathcal{P}$ is the coarsest supporting partition of $\stab_{\consp(G)}(x)$.

  It remains to compute $\orb_{\consp(G)}(x)$. Let $S_0 := \{g\}$ and for each
  $i \geq 0$ let $S_{i+1} := S_i \bigcup_{(u,v) \in \sym_{[n] \setminus \consp(G)}} (u,v)\cdot S_i$. Let  $S$ be the least fixed point. It is easy to see that $S \subseteq \orb_{\consp(G)}(x)$ as every gate in $S$ is equal to the action of some sequence of automorphisms (extending transpositions) acting on $g$. Moreover, since the set of all transpositions generates $\sym_n$, it follows that $\orb_{\consp(G)}(x) \subseteq S$ and so $\orb_{\consp(G)}(x) = S$.

@@

\begin{lem}
  There is an algorithm that takes in a symmetric circuit $C$ of order $n$ with unique labels and outputs for each gate $g$ in $C$ and $a \in \universe{g}$, $\orb_g (a)$ and $\SP_g (a)$.
\end{lem}
\begin{proof}
  Let $(u,v) \in \sym_n$ be a transposition. We can check if $(u,v) \in \stab(g)$ using Lemma~\ref{lem:compute-automorphisms}. Then $(u,v)$ acts like an isomorphism at $g$, and hence $(u,v) \cdot a = L(g)^{-1}((\sigma L(g)(\vec{x}))) (i)$, where $\vec{x} \in \ind(g)$ such that $\vec{x} (i) = a$ for some $i \in [\vert \vec{x} \vert]$. Thus, since Lemma~\ref{lem:compute-automorphisms} gives us that that we can compute $\sigma L(g)(\vec{x})$ in polynomial time, we can compute $(u,v) \cdot a$ in polynomial-time.

  We note that

  and hence we can compute $\sigma \cdot a$ effeciently.  and hence compute $\orb_g(a) = \{ \sigma \cdot a : \sigma \in \stab(g)\}$.

\end{proof}
